### 2024-03-18 supp/supp.qmd
Adding more to supp on unpublised preprints

---
title: "Supplementary Materials for: Empowering the use of community driven preprints in ecology and evolution"
bibliography: ../bib/refs.bib
csl: ../bib/nature.csl
format:
  html:
    toc: true
    toc-location: left
    toc-depth: 3
    toc-title: "**Table of Contents**"
    theme: simplex
    embed-resources: true
    code-fold: show
    code-tools: true
    number-sections: true
    fontsize: "12"
    max-width: "10"
    code-overflow: wrap
  docx:
    reference-doc: ../bib/template.docx
  pdf:
    include-in-header: 
      text: '\usepackage{lineno}\linenumbers'
editor_options: 
  chunk_output_type: console
execute:
  freeze: auto  # re-render only when source changes
  cache: false
  echo: false
  warning: false
  error: false
  include: true
crossref:
  fig-title: 'Figure'
  fig-labels: arabic
  title-delim: "-"
  fig-prefix: "Figure"
  tbl-prefix: "Table"
---

```{r, setup}
#| label: setup

# Load packages
#install.packages("pacman")
pacman::p_load(tidyverse, here, patchwork, readxl, googlesheets4, janitor, maps, forcats,RColorBrewer, flextable, ggimage, gt, performance, ggthemes, mapproj)

# Set rounding options
options(scipen = 1, digits = 2)

# Load in the data
  data <- read_xlsx(here("data", "20231003_EER_preprints_metadata.xlsx"))

# Check duplicate titles. Appears some folks submitted the same paper multiple times instead of simply updating their existing one.

  # First make sure there is no case sensitivity 
  data <- data %>% mutate(title = tolower(`Preprint Title`))

  # Check for duplicates
  dupl <-  data %>%
    group_by(title) %>% 
    summarise(n = n()) %>% 
    filter(n > 1)

  # Remove duplicates
    data <- data %>% 
            distinct(title, .keep_all = TRUE)

  # Check success. Should be 1216. Yes
    dupl2 <- data %>%
            group_by(title) %>% 
            summarise(n = n()) %>% 
            filter(n > 1)

# How many preprints?
  pr <- length(unique(data$`Preprint ID`))

# Dates preprints were published, range and graph
  dates <- range(data$`Published Date`)

# Delegate details
  delegates <- read_sheet("https://docs.google.com/spreadsheets/d/1UEAUZWpOm7C1kKoVoYy-u_D6cbHoF--EGxoY0Gl0qHw/edit#gid=836736319", sheet = "Delegate Details")

  del_details <- delegates  %>% rename("attend" = "Attended hackathon (yes/no)") %>%
			mutate(attend = str_to_title(attend))  %>% tabyl(attend)

# Master preprint list
  master_list <- read_sheet("https://docs.google.com/spreadsheets/d/1UEAUZWpOm7C1kKoVoYy-u_D6cbHoF--EGxoY0Gl0qHw/edit#gid=836736319", sheet = "Master List of Preprints") %>% clean_names()

# Check allocation of papers
		contr_extract <- master_list  %>%  tabyl(extractor_full_name)

	# Check how many each person has done relative to their total
		contr_extract2 <- master_list %>% group_by(extractor_full_name) %>% summarise(n = n(), yes = sum(completed_yes_no_extractor == "Yes"), no = sum(completed_yes_no_extractor == "No"), prop = yes/n)  %>% data.frame()  %>% arrange(n)

# raw data. For analysis
	data_coll <- read_sheet("https://docs.google.com/spreadsheets/d/1032gLryvtCNJ7eJjKBjrn7txxDRUGql9SoIG18QnFrQ/edit?resourcekey#gid=1062856014")

# Final Processed data
  data2 <- read_csv(here("data", "final_data2.csv"))

```

# Materials and Methods

## Pre-registration, Data and Code Availability
We preregistered the study on the Open Science Framework (OSF) (https://osf.io/d7zws) as an open-ended registration. We also created relevant updated study plan releases on our GitHub repository to capture key changes to the project as it developed (https://github.com/daniel1noble/ecoevo_1000). All data and code for reproducing the analyses and conclusions within our paper can be found on GitHub (https://github.com/daniel1noble/ecoevo_1000).

## Data collection
We downloaded metadata on all preprints lodged on *EcoEvoRxiv* between `r gsub("UTC", "", dates[1])` and `r gsub("UTC", "", dates[2])`. This information included data on preprint submission and acceptance dates, date when the preprint was last updated, the reuse license, preprint ID and Digitial Object Identifier (DOI), preprint title, submitting author and co-authors, the number of versions submitted by authors and, if available, a DOI for the published version of the preprint.

As part of the Society for Open, Reliable and Transparent Ecology and Evolutionary Biology's (SORTEE - https://www.sortee.org) conference, we ran a workshop (i.e., Hackathon) where we invited interested SORTEE members and conference registrants to participate in the data collection process. We provided a detailed data collection protocol to all delegates prior to the workshop ([vers. 1.0 - release](https://github.com/daniel1noble/ecoevo_1000/releases/tag/v1.0)). We used the workshop to train participants on how to collect data from *EcoEvoRxiv* preprints. This involved a demonstration of a single preprint followed by data collection by all attendees on the same preprint. Any confusion or questions were addressed by the workshop facilitators and the study data collection instructions and study plan were updated accordingly ([vers.1.1 - release](https://github.com/daniel1noble/ecoevo_1000/releases/tag/V1.1)).

We had a total of `r del_details[2,2]` participants in the workshop, and an additional `r del_details[1,2]` that could not attend but were interested in contributing. Participants assigned themselves to collect data from between `r range(contr_extract$n)[1]` to `r range(contr_extract$n)[2]` preprints (mean = `r mean(contr_extract$n)`, standard deviation = `r sd(contr_extract$n)`). Data collection was achieved using a Google Form to standardise data input. Participants collected the following information for each of their assigned preprints:

- **Preprint DOI**: Copy and pasted from the preprint meta-data file.
- **Submitting/ corresponding authors firstname**: First name (given name) and second name (or initials, if provided) of the author submitting the preprint. Usually there was one submitting author in the meta-data preprint list. If there were multiple authors then the first author listed on the preprint was used.
- **Submitting/ corresponding authors lastname**: Last name of the author submitting the preprint. Usually there was one submitting author in the meta-data preprint list. If there were multiple authors then the first author listed on the preprint was used.
- **Country of the corresponding / submitting author**: Country of affiliation for the author submitting the preprint collecting accoring to standard ISO 3166 names. If teh country could not be determined the affiliation was considered as "NA". 
- **Year of first publication for corresponding/submitted author**: We used [Google Scholar](https://scholar.google.com.au) to collect publication year of the first journal article published by author submitting the preprint. If the submitting author had no profile or did not publish the value enetred was '0'.
- **Taxa being studied**: We collected information about the broad group of taxa that were the focus of the preprint. These categories were "Plants", "Animals", "Fungi", "Algi", "Invertabrates", "Vertebrates", "Microorganisms (bacteria, viruses)", or "Other". Multiple categories could be selected if the preprint contained information on multiple taxa.
- **Discussion on the preprint?**: We identified whether there were community discussions around the preprint on the landing page of each preprint. If there were, we indicated 'yes', otherwise we indicated 'no'. Additional discussions around a preprint may have occurred on other platforms (e.g., Twitter, Facebook, etc.) but we did not collect this information.
- **Type of article**: *EcoEvoRxiv* publishes a greater diversity of preprints compared with other preprint servers. We therefore collected data on article type. The categories were: *Research Article*: any article-like manuscript intended for publication in research journals with new empirical findings; *Methods paper*: papers presenting new methodological or computational approaches; *Reviews and Meta-analyses*: papers quantitatively or qualitatively synthesising a given topic;  *Opinions*: usually short papers providing new perspectives on a topic; *Comments*: papers that explicitly comment on an already published research article; and *Other*: which includes any other category of preprint which may also be government, non-profit, or industry reports, white papers, or other documents that are not intended for publication in a research journal.
- **Link to Data for Preprint**: We were explicitly interested in whether data for the preprint was already publically available. To collect these data we scrolled through the first version of the preprint and any related files on *EcoEvoRxiv* to see if a link (or accession number) to underlying data was provided. If it was reported we took the link to the data.  If the preprint was not an empirical or equivalent article based on data and data analyses (so usually for opinions, comments, narrative reviews, theoretical paper, etc.) we entered 'NA'. If the preprint was based on data but no link (or accession number) to data repository was reported we entered 'none'. If multiple links were provided we took one.
- **Link to Code for Preprint**: We were also interested in whether code for data analysis was already publically available. To collect these data we scrolled through the first version of the preprint and any related files on *EcoEvoRxiv* to see if a link (or accession number) to underlying code was provided. If it was reported we took the link to the code.  If the preprint was not an empirical or equivalent article based on data and data analyses (so usually for opinions, comments, narrative reviews, theoretical paper, etc.) we entered 'NA'. If the preprint was based on data but no link (or accession number) to data repository with code was reported we entered 'none'. If multiple links were provided we took one.
- **Number of citations to preprint**: To understand citation patterns of preprints on *EcoEvoRxiv* we used Google Sscholar to search for the number of citations attributed to the preprint. If the preprint version was not on Google Scholar or was merged with an already published version we considered it as 'NA' (coded as '999'). If the preprint version was online and clearly indicated (i.e., EcoEvoRxiv) then we took the number of citations at the time of collection. While citation counts may vary slightly, the first round of data collection was done in a short time frame. As such, we do not anticipate that citation counts would have changed significantly between the time of data collection and the time of submission of this manuscript.
- **PCI recommendation**: We also deteremined if the preprint been recomended by Peer Community In (PCI). To do this, we searched for any PCI recommendations associated with the preprint on the preprint landing page. If that none was available on the landing page then we searched "peercommunityin.org recommendation "TITLE OF PREPRINT"" on Google. If a link was discovered, we indicated 'yes' and took the DOI link. If not, then we assumed it was not recommended.
- **Publication DOI**: If the preprint had been published as a journal article, we collected the published DOI of that article. 
- **Journal name**: If the preprint had been published as a journal article we recorded the journal name.
- **Publicaton Date**. If the preprint has been published as a journal article, we collected the (first) publication date (Month, Day, Year).
- **Title Change**: If the preprint had been published as a journal article, we assessed whether the title had changed between the first version of the preprint and the published article. Note that any word change was sufficient for a 'yes'. 
- **Number of citations to publication**: We collected the number of citations to the published version of the article manually from [Google Scholar](https://scholar.google.com.au) if published.
- **Link to Data for Publication**: If the preprint was published we collected the link to any published data provided on the publishers webpage or within the published paper. If a link to the data was not visble we recorded the paper as not having provided data ('none'). If the paper had not been published or not based on data and data analyses (opinions, comments, narrative reviews, theoretical paper, etc.) we entered 'NA'. 
- **Link to Code for Publication**: If the preprint was published we collected the link to any published code provided on the publishers webpage or within the published paper. If a link to the data was not visble we recorded the paper as not having provided data ('none'). If the paper had not been published or not based on data and data analyses (opinions, comments, narrative reviews, theoretical paper, etc.) we entered 'NA'.

Additional variables of interest (i.e., journal impact factor, preprint and review policy, gender of submitting author) could be determined oustide of this data collection period.

## Data Checking and Validation
Data from each preprint was cross checked by two authors to ensure accuracy and consistency of the data collected. Any discrepencies were either discussed or fixed prior to analysis.

# Supplementary Results

## Preprint submission and publication dates over time
*EcoEvoRxiv* was established in 2018 and there has been a steady increase in preprints (and postprints) to the journal since that time (@fig-figdates).

```{r, figdates}
#| label: fig-figdates
#| fig-cap: Number of preprints published monthly to *EcoEvoRxiv* between 2018 and 2023.

ggplot(data, aes(x = `Published Date`)) +
  geom_histogram(fill = "grey", color = "black") +
  labs(x = "Published Date", y = "Number of preprints") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

## Publication time of preprints
*EcoEvoRxiv* accepts postprints which means that, technically, a 'preprint' can be published before being lodged on the preprint server. Identifying postprints can be challenging because meta-data does not exist to tag such preprints on *EcoEvoRxiv*. In addition, authors may have lodged a postprint while waiting for an accepted article to be published online for a given research journal.

In many cases, identifying a postprint is straightforward. These are papers that have preprint publication dates that are later than the journal publication date. However, papers that were 'quick' to be accepted could also be postprints, as mentioned above. It seems highly unlikely for a paper to be accepted in a journal and online within a month after being submitted to *EcoEvoRxiv*. We therefore assume that any preprint that was published within a month of being submitted is likely a postprint.

```{r, fig-pubtime}
#| label: fig-pubtime
#| fig-cap: Time between submission and publication of preprints on *EcoEvoRxiv*. 'Preprints' that were published within a month of being submitted are considered postprints.

# How many remain unpublished?
unpub <- data2  %>% filter(is.na(publication_doi))  %>% nrow()

# How many days between submission and publication
preprints <- data2  %>% filter(time_between_preprint_and_pub_days >= 30)
preprints %>% ggplot(aes(x = time_between_preprint_and_pub_days)) + geom_histogram(binwidth = 30) + labs(x = "Days", y = "Frequency") + theme_classic()

# Summary of the data
# what is the range
sum_preprints <- preprints  %>% summarise(min = min(time_between_preprint_and_pub_days), max = max(time_between_preprint_and_pub_days), mean = mean(time_between_preprint_and_pub_days), median = median(time_between_preprint_and_pub_days), sd = sd(time_between_preprint_and_pub_days))
```

Overall, `r unpub` papers were unpublished out of the `r nrow(data2)` (`r (unpub / nrow(data2))*100`%). The maximum time between submission and publication was `r sum_preprints$max` days (@fig-pubtime). The median time to publication was `r sum_preprints$median` days (mean = `r sum_preprints$mean`; SD = `r sum_preprints$sd` days).

## Preprint open access status when published

```{r, oa}

# Summary of the data
sum_is_oa <- tabyl(preprints$is_oa)
sum_oa_status <- tabyl(preprints$oa_status)
```

Preprints can be a useful way to ensure that research remains open and accessible to the research community as the paper is available online no matter what the published status of the preprint ends up being. Most of the published preprints ended up being open access(@fig-openaccess A), however, the status of many (i.e., `r sum_is_oa$n[3]` or `r sum_is_oa$percent[3]*100`%) were unknown, and `r sum_is_oa$n[1]` or `r sum_is_oa$percent[1]*100`% were not published in open access reserach journals. The open access status of the papers that were published in open access journals also varied widely (@fig-openaccess B).

```{r, fig-openaccess}
#| label: fig-openaccess
#| fig-cap: Open access status of preprints when published. A) The number of papers pusblihed in open access journals compared to those not. Note that many journals could not be identified as open access (i.e, NA). B) The classification of open access status of the published preprints.

########-------------------------###
### What is the OA status of the published preprints?
########-------------------------###

# Lets first look only at the preprints. Then, we can look at the postprints. It's possible folks are putting up postprints because they are not OA journals and they want to share their work freely.

p1 <- preprints  %>% ggplot(aes(x = is_oa)) + geom_bar() + labs(x = "Open Access", y = "Frequency") + theme_classic()

p2 <- preprints  %>% ggplot(aes(x = oa_status)) + geom_bar() + labs(x = "Open Access Status", y = "Frequency") + theme_classic()

(p1 + p2) + plot_annotation(tag_levels = 'A', tag_suffix=")")
```

## What countries use preprints the most?

```{r, countries}
#| label: fig-countries
#| fig-cap: Percentage of preprints on *EcoEvoRxiv* across countries.

# Create a global map with countried highlighted and shaded according to the frequency of preprints from that country
# First, we need to get the country names and the number of preprints from each country
  countries <- preprints  %>% tabyl(submitting_author_country)  %>% rename("country" = submitting_author_country, "n" = n)  %>% mutate(country = ifelse(country == "NA", "Unknown", country))  %>% arrange(desc(n))  %>% mutate(country=str_to_title(country))

# Now we can plot the map, but before we do we need to make sure the names in the countries file match those in the region of the world map
  world_map <- map_data(map = "world") %>% 
  filter(! long > 180)

# List unique countries
  maps_country_name <- unique(world_map$region) # Names of regions in maps
  #countries[!countries$country %in% maps_country_name,] # Names of countries in our data that are not in the maps data, we need to fix these in countries

# Fix names in countries
replace = c("United States Of America" = "USA",  "United Kingdom" = "UK",  "Czechia" = "Czech Republic") 
countries <- countries  %>% mutate(country = str_replace_all(country, replace))

ggplot(countries) +
  geom_map(aes(map_id = country, fill = valid_percent*100), map = world_map) +
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group), colour = 'black', fill = NA) +
  expand_limits(x = world_map$long, y = world_map$lat) +
  theme_void() + expand_limits(x = world_map$long, y = world_map$lat) +
  coord_fixed() + labs(fill="%") + theme_map(base_size = 22) + coord_map("moll")

  p1 <- countries  %>% ggplot(aes(map_id = country)) + geom_map(aes(fill = percent), map = map_data("world"), color = "black") + expand_limits(x = map_data("world")$long, y = map_data("world")$lat) + labs(fill = "Number of preprints", title = "Number of preprints by country") + theme_void() + theme(legend.position = "bottom")

# Find the countries with 90% of the preprints

most_use <- countries[cumsum(countries$valid_percent) <= 0.90, ]

```

A total of `r length(unique(countries$country))-1` countries have used *EcoEvoRxiv* as a preprint server. Unsurprisingly, countries in North America and Europe use preprints more exstensively. Countries in Africa, Central America and parts of Asia are still slow to use *EcoEvoRxiv* for their preprints (@fig-countries). Overall, 90% of the preprints come from `r nrow(most_use)-1` countries (@tbl-countries).

```{r}
#| label: tbl-countries
#| tbl-cap: Countries with 90% of the preprints on EcoEvoRxiv.

tbl_most <- most_use %>% filter(!is.na(valid_percent)) %>% mutate(Percentage = valid_percent*100) %>% select(country, Percentage)  %>% rename("Country" = country)

flextable(tbl_most) %>% bold(part = "header") %>% autofit()
```