### 2024-06-07 ms/ms.qmd
Losia's final comments

---
title: "The promise of community-driven preprints in ecology and evolution"
bibliography: ../bib/refs.bib
csl: ../bib/nature.csl
format:
  docx:
    reference-doc: ../bib/template.docx
  pdf:
    include-in-header: 
      text: '\usepackage{lineno}\linenumbers'
editor_options: 
  chunk_output_type: console
execute:
  freeze: auto  # re-render only when source changes
  cache: false
  echo: false
  warning: false
  error: false
  include: true
crossref:
  fig-title: 'Figure'
  fig-labels: arabic
  title-delim: "-"
  fig-prefix: "Figure"
  tbl-prefix: "Table"
---

```{r, setup}
#| label: setup

# Set rounding options
options(scipen = 1, digits = 1)

# Load packages
#install.packages("pacman")
#remotes::install_github("lmullen/genderdata")
pacman::p_load(tidyverse, here, patchwork, readxl, MASS, googlesheets4, janitor, maps, forcats,RColorBrewer, flextable, ggrepel, ggimage, gt, performance, gender,genderdata, ggthemes, mapproj, DemografixeR, countrycode, see, glmmTMB, cowplot, viridis, cowplot, magick, lme4, lmerTest, data.table)

# Load in the data
  data <- read_xlsx(here("data", "20231003_EER_preprints_metadata.xlsx"))

# Check duplicate titles. Appears some folks submitted the same paper multiple times instead of simply updating their existing one.

  # First make sure there is no case sensitivity 
  data <- data %>% mutate(title = tolower(`Preprint Title`))

  # Check for duplicates
  dupl <-  data %>%
    group_by(title) %>% 
    summarise(n = n()) %>% 
    filter(n > 1)

  # Remove duplicates
    data <- data %>% 
            distinct(title, .keep_all = TRUE)

  # Check success. Should be 1216. Yes
    dupl2 <- data %>%
            group_by(title) %>% 
            summarise(n = n()) %>% 
            filter(n > 1)

# How many preprints?
  pr <- length(unique(data$`Preprint ID`))

# Dates preprints were published, range and graph
  dates <- range(data$`Published Date`)

# Delegate details
  #delegates <- read_sheet("https://docs.google.com/spreadsheets/d/1UEAUZWpOm7C1kKoVoYy-u_D6cbHoF--EGxoY0Gl0qHw/edit#gid=836736319", sheet = "Delegate Details")

  #del_details <- delegates  %>% rename("attend" = "Attended hackathon (yes/no)") %>%
			#mutate(attend = str_to_title(attend))  %>% tabyl(attend)

# Master preprint list
  #master_list <- read_sheet("https://docs.google.com/spreadsheets/d/1UEAUZWpOm7C1kKoVoYy-u_D6cbHoF--EGxoY0Gl0qHw/edit#gid=836736319", sheet = "Master List of Preprints") %>% clean_names()

# Check allocation of papers
		#contr_extract <- master_list  %>%  tabyl(extractor_full_name)

	# Check how many each person has done relative to their total
		#contr_extract2 <- master_list %>% group_by(extractor_full_name) %>% summarise(n = n(), yes = sum(completed_yes_no_extractor == "Yes"), no = sum(completed_yes_no_extractor == "No"), prop = yes/n)  %>% data.frame()  %>% arrange(n)

# raw data. For analysis
	#data_coll <- read_sheet("https://docs.google.com/spreadsheets/d/1032gLryvtCNJ7eJjKBjrn7txxDRUGql9SoIG18QnFrQ/edit?resourcekey#gid=1062856014")

# Final Processed data
  data2 <- read_csv(here("data", "final_data2.csv"))

# Subset the relevant rows
		data3  <- data  %>%  dplyr::select(`Preprint ID`, `Preprint DOI`, `Published Date`)  %>% rename("preprint_pub_date" = "Published Date",
                                                                                                      "preprint_id"  = "Preprint ID")

# Merge with the main data to make sure that the published data of the preprint is known for ALL papers
    data2  <- left_join(data2, data3, by = c("preprint_id" = "preprint_id"))

```

Authors: Daniel W.A. Noble ^1^, $\ddagger$, Zoe A. Xirocostas, Nicholas C. Wu, April R. Martinig, Shinichi Nakagawa^2,\*^,Malgorzata Lagisz^2,\*^

Affiliations:

^1^ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2600, Australia &NewLine;  
^2^ Ecology and Evolution Research Centre, School of Biological, Earth and Environmental Sciences, University of New South Wales, Sydney, NSW, Australia &NewLine;

$\ddagger$ corresponding author, daniel.noble@anu.edu.au &NewLine;

\* equal contribution

# Introduction

Publishing preprints -- papers communicating non-peer-reviewed research findings -- is now an entrenched practice across a multitude of scientific disciplines [@Ginsparg2011]. Preprints in biology have had a slower uptake relative to other disciplines [@Berg2016], but new discipline-specific preprint servers, such as *EcoEvoRxiv* (https://ecoevorxiv.org), provide a means by which ecologists and evolutionary biologists can disseminate research findings. Preprints attempt to break down barriers to scientific publishing by: 1) increasing the visibility of research and the speed at which research findings become available, which can lead to more citations [e.g., @colavizza2024analysis; @fu2019releasing]; 2) helping establish the precedence of research findings; 3) removing financial barriers to open access publication; and 4) enabling feedback from the research community [@Proulx2013; @bourne2017ten; @vale2015accelerating]. Ultimately, preprints can facilitate the rapid sharing of scientific knowledge that can have significant impacts on fundamental and applied knowledge globally [@ni2024preprint].

Preprint servers can empower researchers to make their research findings more accessible, open, and transparent but only if they are used as forums for spreading and discussing findings within a research community. However, significant barriers to the widespread adoption of preprints remain, ranging from a lack of clarity around preprint policies in journals [@Klebel2020] to a stigma within the research community that preprints are of poor quality [@Chiarelli2019] (but see [@Faser2022]). Nonetheless, we lack an understanding of the factors that influence preprint use in ecology and evolution. Such an understanding may help improve current initiatives (see below), inform future ones and allow us to work harder in further breaking down barriers to scientific publishing.

*EcoEvoRxiv* is one of the few community-driven preprint servers that has paved the way for new initiatives, by accepting multilingual preprints, registered reports, and non-traditional research reports. *EcoEvoRxiv* also promotes peer review and community discussion in the hopes of improving the quality of preprints and speeding up their peer-reviewed publication (e.g., Peer Community In -- PCI [@guillemaud2019peer]). Peer Community In allows for fast, constructive peer review around a preprint with peer reviews being transparent and published online [@guillemaud2019peer].

Here, we explore the first preprints uploaded to *EcoEvoRxiv* to characterise preprint practices in ecology and evolution. We aim to understand: 1) in what countries authors who use *EcoEvoRxiv* are located; 2) the taxonomic diversity study systems across preprints; 3) whether preprint server use depends on career stage and gender; 4) the extent to which authors make use of preprint servers for reports and community-driven peer review; 5) the extent to which data and code are shared in preprints; and 6) how many preprints remain unpublished, and for those that are published, how long it took for them to become published. In the process, we also provide a summary of what makes *EcoEvoRxiv* distinct from other preprint servers to help further clarify the benefits of using community-driven preprint servers to disseminate research findings.

# Getting to know your *EcoEvoRxiv* preprint server

*EcoEvoRxiv* is run by the Society for Open, Reliable, and Transparent Ecology and Evolutionary Biology (SORTEE)[@o2021towards]. Originally launched in 2018 on the Center for Open Science preprint platform, *EcoEvoRxiv* has become a popular preprint server for ecologists and evolutionary biologists. The server has since been adopted by the California Digital Library (CDL). Editors are ecologists and evolutionary biologists from across the globe who volunteer their time to screen papers and push new initiatives in the preprint space. *EcoEvoRxiv* allows authors to post both preprints and postprints (also known as author-accepted manuscripts). While preprints are versions of manuscripts posted by authors before peer-review, postprints are versions of peer-reviewed and accepted articles but without typesetting and formatting by a journal. The main reason for publishing postprints on a preprint server is to ensure published articles are openly accessible to everyone without a paywall (i.e., green open access). Postprints can be published anytime, provided that journals allow it (which most do; see https://www.sherpa.ac.uk/romeo/).

```{r, fig-summary}
#| label: fig-summary
#| fig-cap: Summary of articles posted to *EcoEvoRxiv*. A) Number of articles (preprints and postprints) published on *EcoEvoRxiv* between 2018 and 2023. *EcoEvoRxiv* was established in June 2018 before the launch in November 2018. Notable milestones include *EcoEvoRxiv* transitioning to the California Digital Library (CDL), the acceptance of preprints and postprints in Spanish and Portuguese, and the acceptance of the first IUCN Red List Ecosystem report; B) Geographic origin of the preprints and postprints uploaded to *EcoEvoRxiv*, inferred from the country of affiliation of the submitting author; C) Taxa used/covered in the articles posted to *EcoEvoRxiv* (n = 1080 articles); D) Types of preprints accepted on *EcoEvoRxiv* (n = 620 articles). E) Academic age of authors posting to *EcoEvoRxiv* along with the gender of the submitting author. Values lower than zero are indicative of authors who uploaded preprints before their first scientific publication in a journal. Map base source: R Package "maps" v.3.4.2. Shapefile: Natural Earth https://www.naturalearthdata.com/about/terms-of-use/.

#| echo: false
#| warning: false
#| message: false

  # Subset the preprints
  preprints <- data2  %>% filter(time_between_preprint_and_pub_days >= 60)

  dates_preprints <- range(preprints$preprint_published_date)
##---------------------##
# Articles on EcoEvoRxiv over time
##---------------------##
  angle = 45
  p1 <- ggplot(data2, aes(x = preprint_pub_date)) +
    geom_histogram(fill = "white", color = "black") +
    scale_y_continuous(limits = c(0, 130), expand = c(0, 0)) +
    labs(x = "Published date", y = "Number of papers") +
    theme_classic() +
    #theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    theme(plot.tag = element_text(size = 16, face = "bold")) + 
      # Soft Launch of EcoEvoRxiv
    annotate( "segment", x = as.POSIXct("2018-06-01"),  xend = as.POSIXct("2018-06-01"), y = 0, yend = 82, color = "black", linetype = "dashed", size = 0.5) + 
    annotate("text", x = as.POSIXct("2018-06-01"), y = 85, label = "Soft Launch", size = 4, angle = angle, hjust = 0) +
    # Launch of EcoEvoRxiv
    annotate( "segment", x = as.POSIXct("2018-11-11"),  xend = as.POSIXct("2018-11-11"), y = 0, yend = 82, color = "black", linetype = "dashed", size = 0.5) + 
    annotate("text", x = as.POSIXct("2018-11-11"), y = 85, label = "Launch", size = 4, angle = angle, hjust = 0) +
    # Spanish and Port
    annotate( "segment", x = as.POSIXct("2023-02-14"),  xend = as.POSIXct("2023-02-14"), y = 0, yend = 82, color = "black", linetype = "dashed", size = 0.5) + 
    annotate("text", x = as.POSIXct("2022-11-14"), y = 85, label = "Spanish and Portug.", size = 4, angle = angle, hjust = 0) +
    # Transition to CDL
    annotate( "segment", x = as.POSIXct("2022-09-14"),  xend = as.POSIXct("2022-09-14"), y = 0, yend = 82, color = "black", linetype = "dashed", size = 0.5) + 
    annotate("text", x = as.POSIXct("2022-07-14"), y = 85, label = "CDL", size = 4, angle = angle, hjust = 0) +
   # IUCN Red Eco submission 29 August 2023
   # 1000 preprints
    annotate( "segment", x = as.POSIXct("2023-08-29"),  xend = as.POSIXct("2023-08-29"), y = 0, yend = 82, color = "black", linetype = "dashed", size = 0.5) + 
    annotate("text", x = as.POSIXct("2023-04-29"), y = 85, label = "IUCN Report", size = 4, angle = angle, hjust = 0)

  ggsave(here("output", "figs", "published_date.svg"), p1, width = 7.5, height = 2)

  #Plot was then imported to illustrator for aesthetic touches and to combine with other plots in Figure 1. - Zoe X
##---------------------##
# Countries using EcoEvoRxiv
##---------------------##
   # Create a global map with countried highlighted and shaded according to the frequency of preprints from that country
  # First, we need to get the country names and the number of preprints from each country
    countries <- data2  %>% tabyl(submitting_author_country)  %>% rename("country" = submitting_author_country, "n" = n)  %>% mutate(country = ifelse(country == "NA", "Unknown", country))  %>% arrange(desc(n))  %>% mutate(country=str_to_title(country))

  # Now we can plot the map, but before we do we need to make sure the names in the countries file match those in the region of the world map
    world_map <- map_data(map = "world") %>% 
    filter(! long > 180)

  # List unique countries
    maps_country_name <- unique(world_map$region) # Names of regions in maps
    #countries[!countries$country %in% maps_country_name,] # Names of countries in our data that are not in the maps data, we need to fix these in countries

  # Fix names in countries
      replace <- c("United States Of America" = "USA",  "United Kingdom" = "UK",  "Czechia" = "Czech Republic") 
    countries <- countries  %>% mutate(country = str_replace_all(country, replace))  %>% filter(!is.na(valid_percent ))  %>% mutate(valid_percent_dis = factor(round(valid_percent, 3)*100, ordered = TRUE))

    map_fig <- ggplot(countries) +
    geom_map(aes(map_id = country, fill = valid_percent_dis), map = world_map) +
    geom_polygon(data = world_map, aes(x = long, y = lat, group = group), colour = 'black', fill = NA) +
    expand_limits(x = world_map$long, y = world_map$lat) +
    theme_void() + expand_limits(x = world_map$long, y = world_map$lat) + scale_fill_viridis(option = "G", discrete = TRUE, direction = -1) +
    coord_fixed() + labs(fill="Percentage (%) of Total") + theme(plot.tag = element_text(size = 16, face = "bold"))

  ggsave(here("output", "figs", "worldmap.svg"), map_fig, width = 15, height = 20)
  #removed Antarctica and sub-antarctic islands in Illustrator and combined with other plots to create Figure 1. -Zoe X

  # Find the countries with 90% of the preprints

  most_use <- countries[cumsum(countries$valid_percent) <= 0.90, ]

##---------------------##
# Taxa
##---------------------##
  cols <- c("plants", "algi", "fungi", "microorganisms", "invertebrates", "vertebrates")

  # Summary of the data
  sum_taxa <- data2  %>% dplyr::select(all_of(cols))  %>% apply(., 2, sum, na.rm = TRUE)  %>% data.frame()   %>% rownames_to_column(var = "taxa")  %>% rename("n" = ".")  %>% mutate(perc = n/sum(n)*100, labels = paste0(round(perc, 1), "%"))
  #sum(sum_taxa$n) #1080
  # Use data to create a pir chart and clean up colour palette

  # Use data to create a pie chart and clean up colour palette
  taxa_fig <- ggplot(sum_taxa, aes(x = fct_reorder(taxa, perc), y = perc)) +
    geom_col(color = "black", fill = "white") +
    scale_y_continuous(expand = c(0, 0)) +
    #scale_fill_viridis_d() +
    #coord_polar(theta = "y") + 
    theme_classic() + labs(y = "Percentage (%) of preprints", x = "Studied taxa") + 
    theme(plot.tag = element_text(size = 16, face = "bold")) + coord_flip() +
    scale_fill_colorhex_d(palette = 1042984, discrete = TRUE)

##---------------------##
# Types of preprints
##---------------------##
# Summary of the data
  sum_types <- preprints  %>% tabyl(type_of_preprint)  %>% rename("type" = type_of_preprint, "n" = n)  %>% mutate(perc = n/sum(n)*100, labels = paste0(round(perc, 1), "%"))
  #sum(sum_types$n) #620

  type_fig <- ggplot(sum_types, aes(x = fct_reorder(type, perc), y = perc)) +
    geom_col(color = "black", fill = "white") +
    scale_y_continuous(expand = c(0, 0)) +
    #scale_fill_viridis_d() +
    #coord_polar(theta = "y") + 
    theme_classic() + labs(y = "Percentage (%) of preprints", x = "Article type") + 
    theme(plot.tag = element_text(size = 16, face = "bold")) + coord_flip() +
    scale_fill_colorhex_d(palette = 1042984, discrete = TRUE) + theme(legend.position = "none")

type_taxa <- cowplot::plot_grid(taxa_fig, 
                   type_fig, 
                   align = "v", ncol = 2, labels = c('c', 'd'))

  ggsave(here("output", "figs", "taxa_type.svg"), type_taxa)

  #Plot was then imported to illustrator for aesthetic touches and to combine with other plots in Figure 1. - Zoe X

nt <- sum_types  %>% filter(type %in% c("book", "report", "book chapter", "other"))  %>% summarise(perc = sum(perc))

##-------------------##
# Career Stage & Gender
##-------------------##

#compute academic age! (Zoe X)
preprints <- preprints %>% 
  mutate(preprint_published_year_z = year(preprint_published_date)) %>% 
  mutate(academic_age = (preprint_published_year_z-submitting_author_first_publication_year))

academic_age <- ggplot(preprints, aes(x = academic_age)) +
  geom_histogram(binwidth = 1, fill = "white", color = "black") +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "Years since submitting author's first publication (academic age)", y = "Number of preprints") +
  geom_vline(aes(xintercept = median(academic_age)), color = "black", linetype = "dashed", size = 0.5) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#mean(age$academic_age)

ggsave(here("output", "figs", "academic_age.svg"), academic_age, width = 7.5, height = 2)

  #Plot was then imported to illustrator for aesthetic touches and to combine with other plots in Figure 1. - Zoe X

count_dat  <- preprints  %>% 
              distinct(submitting_author, academic_age) %>% 
              group_by(academic_age) %>% 
              summarise(n = n())

model <- glm.nb(n ~ academic_age, data = count_dat)
res <- summary(model)$coefficients

# First need to get the authors first name
  data2 <- data2 %>% mutate(submitting_author_firstname = str_split(submitting_author, " ") %>% map_chr(1))

# In order to localise by country we need to get the country code
  data2 <- data2 %>% mutate(submitting_author_country_code = countrycode(data2$submitting_author_country, origin = "country.name", destination = "iso2c"))

# Lets have a quick look at whether these are 'correct'
  #data2 %>% dplyr::select(submitting_author_country_code, submitting_author_country)  %>% distinct()

# This looks good, but lets see which two are NA. These are in fact NA, even though they shouldn't really be as affliations are not listed on preprint. Even the published paper isn't clear. 
  #data2 %>% filter(is.na(submitting_author_country_code))  %>% data.frame()

# Get gender using all unique first names from the data set
data_genders <- data.frame(gender(unique(data2$submitting_author_firstname)))
#dim(data_genders) #514 assigned
#names(data_genders)
colnames(data_genders)[1]  <- c("submitting_author_firstname")

# Merge with main data frame
data3 <- left_join(data2, data_genders, by = c("submitting_author_firstname"))
#names(data3)
#table(is.na(data3$proportion_male)) #157 missing

# Pull out records missing names
#data3[is.na(data3$proportion_male), c("submitting_author", "submitting_author_country") ] #157 missing
#data3[is.na(data3$proportion_male) == FALSE & data3$proportion_male > 0.05 & data3$proportion_male < 0.95, c("submitting_author", "submitting_author_country") ] # 84 missing a bit unsure

#transliterate the non-ASCII characters using stringi package, e.g.: #stri_trans_general("Zażółć gęślą jaźń", "Latin-ASCII")
data3$submitting_author_ASCII <- stringi::stri_trans_general(data3$submitting_author, "Latin-ASCII")

#data3 %>% 
  #filter(is.na(data3$proportion_male) == TRUE | (proportion_male > 0.05 & proportion_male < 0.95)) %>% dplyr::select(submitting_author_ASCII, submitting_author, submitting_author_country, preprint_title, gender) %>% 
 # write.csv(file = here("data", "final_data3_fillin.csv")) #used to manually check and fill in gender data - then saved as final_data3_fillin_ML.csv

# Load updated data	subset
data3_update <- read_csv(here("data", "final_data3_fillin_ML.csv"))
#dim(data3_update)
#names(data3_update)

data4 <- left_join(data3, data3_update, by = c("submitting_author_ASCII", "submitting_author", "submitting_author_country", "preprint_title"))
#table(data4$gender.x, useNA = "always")
#table(data4$gender.y, useNA = "always")
data4$gender.x[is.na(data4$gender.x)] <- data4$gender.y[is.na(data4$gender.x)] #replace with correct values
data4 <- rename(data4, gender = gender.x)
#table(data4$gender, useNA = "always") #more males than females, two NA ue to missing first name / collective authorship

# Plot the data
###Zoe's gender pie###
#female - 37.8, male - 62.2

gender <- c("female", "male")
percentage <- c(37.8, 62.2)
sex_ratio<- data.frame(gender, percentage)

sex_ratio <- sex_ratio %>% ggplot(aes(x = "", y = percentage, fill = gender)) +
    geom_col(color = "black")+
  scale_fill_manual(values = c("white", "black")) +
    coord_polar(theta = "y") + 
    theme_void()

ggsave(here("output", "figs", "sex_ratio.svg"), sex_ratio)

##Pie was then imported to illustrator for aesthetic touches and to combine with other plots in Figure 1. - Zoe X

# Load in and plot the final figure
  fig <- image_read(here("output", "figs", "Figure 1_FINAL.png")) 
  fig

```

### *Overview of **EcoEvoRxiv** preprints (and postprints)*

To better understand preprint (and postprint) use on *EcoEvoRxiv*, we downloaded metadata on the articles available on *EcoEvoRxiv* as of `r gsub("UTC", "", dates[2])` (see Supplement for more details on methods). We consider both preprints and postprints as ‘articles’. After removing five duplicate titles – suggesting that a few authors created multiple submissions for the same preprint rather than updating the existing submission – we had data for a total of `r pr` articles with ~55--60 preprints published approximately monthly in the last two years ([@fig-summary]A).  *EcoEvoRxiv* hosts articles from authors based in `r length(unique(countries$country))` countries, with 90% of the articles coming from just `r nrow(most_use)-1` countries. North America, Australia, and European countries upload the most preprints, with many fewer coming from countries in Africa, Central America, and parts of Asia ([@fig-summary]B).  Articles covered all major taxonomic groups, with the most common being vertebrates (`r sum_taxa$perc[6]`%), plants (`r sum_taxa$perc[1]`%), and invertebrates (`r sum_taxa$perc[5]`%) ([@fig-summary]C).

### *Diversifying article types on **EcoEvoRxiv**: overcoming the 'grey literature' problem*

Accepting a greater diversity of article types allows *EcoEvoRxiv* to help deal with the 'grey literature' problem, whereby data that are relevant for research syntheses are not published in typical peer-reviewed journals [@haddaway2015shades; @haddaway2020eight]. *EcoEvoRxiv* has made a concerted effort to diversify the types of articles accepted. This is reflected by `r nt`% of the articles on *EcoEvoRxiv* being books, book chapters, reports, and other research output types, which are typically considered ‘grey literature’ in ecology and evolutionary biology. As a result, articles on *EcoEvoRxiv* are more diverse than those on preprint servers which have more restrictive submission policies. For example, *bioRxiv* only accepts research articles (https://www.biorxiv.org/submit-a-manuscript).

Typical research articles are still the most common type of preprint on *EcoEvoRxiv* (`r sum_types[sum_types$type == "research article",]$perc`%), followed by reviews and meta-analyses (`r sum_types[sum_types$type == "reviews and meta-analyses",]$perc`%) and opinion papers (`r sum_types[sum_types$type == "opinion",]$perc`%) ([@fig-summary]D). Currently, *EcoEvoRxiv* does not host many reports, particularly from government or industry, but has formed fruitful partnerships with the International Union for Conservation of Nature (IUCN). For example, IUCN Red-list Ecosystem Reports are now posted to *EcoEvoRxiv* and our community has been able to work closely with the IUCN to ensure these documents meet the IUCN requirements. We would encourage authors to consider posting books, book chapters, and reports to ensure that they are openly accessible and more easily found. *EcoEvoRxiv* articles are given a unique DOI and are indexed on Google Scholar.

### *Breaking down language barriers to scientific communication: improving diversity and data representation globally*

A significant barrier to the communication of research findings is the fact that they are primarily communicated in English [@amano2016languages; @amano2013four; @amano2021ten]. Research communication through a single language has major consequences for the global distribution of knowledge, resulting in knowledge gaps across some of the most biodiverse and threatened regions in the world [@amano2023role; @zenni2023multi]. Such gaps also impact research syntheses and meta-analyses because they create a distorted picture of our knowledge base that can affect future research, policy development and decision-making [@hannah2024language; @zenni2023multi; @white2021geographical; @konno2020ignoring].

*EcoEvoRxiv* is the only preprint server to date that breaks down language barriers to scientific communication by accepting not only English, but also Spanish, Portuguese and French language articles. *EcoEvoRxiv* plans to expand to other languages as new non-English editors become available. Such initiatives are incredibly important if we are to begin filling global voids of scientific knowledge. However, multilingual initiatives have been slow to take off on *EcoEvoRxiv*, with only a few Spanish articles, and a single Portuguese article, posted since starting to accept non-English articles in 2023. Part of the challenge in getting authors to submit non-English articles is the lack of awareness of *EcoEvoRxiv* in non-English speaking countries, cultural differences in the perception of preprints, and a strong reliance on traditional publishing models that typically mandate publishing in English [@arenas2024academic].

### *Generational and gender-based gaps in preprinting practices*

Research papers can take a while to be published (see below). However, Early and Mid-Career Researchers (EMCRs) (~10 years post-PhD) are under pressure to publish rapidly to be competitive in job applications, promotions, and obtaining grants to progress their careers [@vale2015accelerating; @sarabipour2019value]. Preprints are one way EMCRs can achieve faster dissemination and greater visibility [@fu2019releasing]. As such, EMCRs may be expected to make use of preprints more than colleagues at later career stages. We collected data on the 'academic age' of submitting authors by looking at Google Scholar profiles of authors (when available) and recording their first year of publication in a peer-reviewed journal. While this is a rough estimate of career stage, there was evidence that the number of preprints posted decreases with later career stages (negative binomial glm: year slope = `r res[2, "Estimate"]`, SE: `r round(res[2, "Std. Error"], 2)`, *p* < 0.001, n = `r nrow(count_dat[complete.cases(count_dat),])` years). Most preprints were submitted by authors who published their first paper in the last ~10 years ([@fig-summary]E), with the median year since first publication being `r median(data2$submitting_author_first_publication_year, na.rm = TRUE)` (mean = `r mean(data2$submitting_author_first_publication_year, na.rm = TRUE)`; SD = `r sd(data2$submitting_author_first_publication_year, na.rm = TRUE)`, n = `r sum(complete.cases(data2$submitting_author_first_publication_year))`). These patterns support the expectation that EMCRs may use preprints to make their work more visible and disseminate their findings more quickly. However, we acknowledge that validating this conclusion does require more rigorous experimental approaches.

Gender differences in preprint use and publication outcomes have also been observed in several research fields, including ecology and evolutionary biology [@fox2019gender; @wehner2020comparison]. Therefore, such discrepancies are expected to manifest in preprint use on *EcoEvoRxiv*, but it is unclear to what extent. Understanding gender publishing patterns is challenging with observational data such as ours because we cannot know the gender of authors for certain, but we can use a data-driven approach to ascertain the probability that a particular name is of a given gender (man or woman). To obtain a rough idea of an author's gender, we used the R package *gender* (v.`r utils::packageVersion("gender")`; @mullen2021predict) to predict the most likely gender of the submitting author of a preprint. We only used algorithm-assigned names where the gender was identified with 95% certainty. For the remaining names, we performed manual searches to determine gender based on the pronouns and photographs from professional and personal websites. We acknowledge that our approach does not capture self-assigned and non-binary genders. As such, our assumptions about an author's gender identity may be incorrect. Our data on gender had only two missing values--one where the first name of the submitting author was missing and the other one for a collective submission. As expected, we found that women were less likely to publish on *EcoEvoRxiv* compared to men (women: `r (table(data4$gender)[1] / sum(table(data4$gender))*100)`%), which may reflect the broader publishing disparities between male and female scientists [@fox2019gender].

# Following the journey of a preprint on *EcoEvoRxiv*: from submission to publication
```{r, fig-pubsummary}
#| label: fig-pubsummary
#| echo: false
#| warning: false
#| message: false
#| fig-cap: Summary of the publication status of articles on *EcoEvoRxiv*. A) Time between uploading an preprint to *EcoEvoRxiv* and its publication as a peer-reviewed journal article. Articles that were published within 2 months (60 days) of being submitted are considered postprints and are not included in this figure. B) Access status of articles linked to *EcoEvoRxiv* preprints classified as “Open Access” or “Not Open Access”. “Unknown” status is for articles not explicitly identified as being open access or not on the Unpaywall platform. C) Sub-types of open access status of articles linked to *EcoEvoRxiv* preprints. For full details on the meaning of each category see  https://support.unpaywall.org/support/solutions/articles/44001777288-what-do-the-types-of-oa-status-green-gold-hybrid-and-bronze-mean-.

##---------------------##
# Publication times
##---------------------##

# How many remain unpublished?
unpub <- data2  %>% filter(is.na(publication_doi))  %>% nrow()

# Summary of the data
# what is the range
sum_preprints <- preprints  %>% summarise(min = min(time_between_preprint_and_pub_days), max = max(time_between_preprint_and_pub_days), mean = mean(time_between_preprint_and_pub_days), median = median(time_between_preprint_and_pub_days), sd = sd(time_between_preprint_and_pub_days))

p1_pub_zoe <- preprints %>% ggplot(aes(x = time_between_preprint_and_pub_days)) +
  geom_histogram(binwidth = 30, fill = "white", color = "black") +
  theme_classic() +  geom_vline(aes(xintercept = sum_preprints$median), colour = "black", linetype = "dashed", linewidth = 0.5) + annotate("text", x = sum_preprints$median + 650, y = 60, label = paste0(sum_preprints$median, " days", "(~ ", sum_preprints$median/30, " months)"), colour = "black") + labs(x = "Time to publication (days)", y = "Number of preprints") + theme_classic() + theme(axis.title = element_text(size = 10)) +
  scale_y_continuous(limits = c(0, 80), expand = c(0, 0))

ggsave(here("output", "figs", "time_to_publication.svg"), p1_pub_zoe)

#Plot was then imported to illustrator for aesthetic touches and to combine with other plots in Figure 2. - Zoe X

##-------------------##
# Open access
##-------------------##

# Summary of the data
sum_is_oa <- tabyl(preprints$is_oa)
sum_oa_status <- tabyl(preprints$oa_status)

p2_pub <- preprints  %>% ggplot(aes(x = is_oa)) + geom_bar() + labs(x = "Open Access", y = "Number of papers") + theme_classic()  + scale_y_continuous(expand = c(0, 0))

ggsave(here("output", "figs", "Open_access_TRUE_FALSE.svg"), p2_pub)

#Plot was then imported to illustrator for aesthetic touches and to combine with other plots in Figure 2. - Zoe X

p3_pub <- preprints %>% mutate(oa_status = (fct_relevel(oa_status, c("gold","hybrid", "bronze", "green", "closed", "NA"))))  %>% ggplot(aes(x = oa_status, fill = oa_status)) + geom_bar() + labs(x = "Open Access type", y = "Number of papers") + theme_classic() + theme(axis.title = element_text(size = 14), legend.position = "none") + scale_fill_manual(values = c("bronze" = "#9b6023", "gold" = "#ffd700", "green" = "#008000", "hybrid" = "#6fa8dc", "NA" = "grey", "closed" = "black"))+ scale_y_continuous(expand = c(0, 0))

ggsave(here("output", "figs", "Open_Access_type.svg"), p3_pub)

#Plot was then imported to illustrator for aesthetic touches and to combine with other plots in Figure 2. - Zoe X

##-------------------##
# Data and Code Sharing
##-------------------##

# Summary of the data
  preprint_data <- preprints  %>% tabyl(data_link_preprint_cleaned) 
  preprint_code <- preprints  %>% tabyl(code_link_preprint_cleaned) 
  article_data <- preprints  %>% tabyl(data_link_article_cleaned) 
  article_code <- preprints  %>% tabyl(code_link_article_cleaned)

# Check if it was shared not in preprint but then in article
  pre_article_data <- preprints   %>%  filter(!data_link_preprint_cleaned == "not applicable") %>% mutate(data_shared = ifelse(data_link_preprint_cleaned == "no" & data_link_article_cleaned == "yes", "yes", ifelse(code_link_preprint_cleaned == "yes" & code_link_article_cleaned == "yes", "always", "no")))  %>% tabyl(data_shared)
  pre_article_code <- preprints  %>%  filter(!data_link_preprint_cleaned == "not applicable") %>% mutate(code_shared = ifelse(code_link_preprint_cleaned == "no" & code_link_article_cleaned == "yes", "yes", ifelse(code_link_preprint_cleaned == "yes" & code_link_article_cleaned == "yes", "always", "no")))  %>% tabyl(code_shared)

##-------------------##
# PCI Recommendations
##-------------------##
  # How many articles were PCI recomended
  pci <- preprints  %>% tabyl(pci_recommendation_preprint)

##-------------------##
# Title changes
##-------------------##
titles <- preprints %>%  filter(!is.na(publication_doi)) %>% tabyl(publication_title_changed)

fig2 <- image_read(here("output", "figs", "Figure 2_final.png"))
fig2
```

### *Science takes time, but publication could take longer*
Increased competition in science has raised the bar with respect to the amount of data required for publication [@vale2015accelerating]. This requirement is a good outcome if it results in higher-impact research that better clarifies our understanding of the natural world, but it does come at a cost for the speed of research dissemination [@vale2015accelerating]. Preprints have been proposed as a way to disseminate research more quickly as it can take a long time before results are ultimately published [@bourne2017ten; @vale2015accelerating]. However, data on the time to publication is needed to quantify the real benefit of preprints in this context.

We estimated how long it takes to publish a research paper in ecology and evolution by recording the time between when an article was first posted on *EcoEvoRxiv*, and its final acceptance in a peer-reviewed journal. In total, `r unpub` papers remained unpublished (`r (unpub / nrow(data2))*100`%, n = `r nrow(data2)`) at the time when these data were collected. Not all of these papers, however, are anticipated to be published in a peer-reviewed journal (e.g., reports). Nonetheless, the median time to publication for preprints was `r sum_preprints$median` days (8 months) (mean = `r sum_preprints$mean`; SD = `r sum_preprints$sd` days) with the maximum time to publication being `r sum_preprints$max` days or `r sum_preprints$max/365` years ([@fig-pubsummary]A). Our results largely confirm the extended timeframes that most authors experience between writing their research papers and their publication.

### *Cautious 'open'-mindedness of research in preprints*
In addition to speeding up dissemination, preprints and postprints can also be a useful way to ensure that research remains open and accessible to the research community irrespective of the accessibility of the final peer-reviewed paper [@bourne2017ten; @vale2015accelerating]. We evaluated whether articles hosted at *EcoEvoRxiv*, and that were also published in a journal, were published open access. The open access status of each published article was obtained using the R package *roadoi* (v.`r utils::packageVersion("roadoi")`) to connect to the Unpaywall platform [@jahn2024]. Most of the published articles were open access [`r sum_is_oa$valid_percent[2]*100`% (*n* = `r sum_is_oa$n[2]` out of `r sum(sum_is_oa$n[1:2])` where the status was known); [@fig-pubsummary]A]; however, `r sum_is_oa$valid_percent[1]*100`% (*n* = `r sum_is_oa$n[1]`) were published behind paywalls. For articles published in open access journals, the type of open-access also varied widely (e.g., Gold, Hybrid, Green OA etc., [@fig-pubsummary]B).

Data and code sharing are also key components of open science [@roche2015public]. In the spirit of 'openness', we expected data and code sharing among preprints to be greater than in many papers published in research journals [@roche2015public; @o2021preferred]. Despite this, we found that `r (preprint_data[1,"n"]/sum(preprint_data[c(1,3),"n"]))*100`%  (*n* = `r preprint_data[1,2]`) of articles relying on data on *EcoEvoRxiv* did not share data, and `r (preprint_code[1,"n"]/sum(preprint_code[c(1,3),"n"]))*100`% (*n* = `r preprint_code[1,2]`) did not share code (counting only data-based articles, i.e excluding reviews, commentaries or theoretical works).

Authors may be reluctant to share data and code for preprints because of the perceived concern that others may acquire and use their data and code before publication in a journal. Authors of `r pre_article_data[3,"percent"]*100`% (*n* = `r pre_article_data[3,2]`) of articles that did not share data at the preprint stage did ultimately share data when the article was published; whereas authors of `r pre_article_data[2,"percent"]*100`% (*n* = `r pre_article_data[2,2]`) never shared data. However, `r pre_article_data[1,"percent"]*100`% (*n* = `r pre_article_data[1,2]`) shared data at both stages. The same was true for code. Overall, `r pre_article_code[3,"percent"]*100`% (*n* = `r pre_article_code[3,2]`) preprints had no open code at the preprint stage but did at the published article stage and authors of `r pre_article_code[2,"percent"]*100`% (*n* = `r pre_article_code[2,2]`) preprints did not share code at either stage. However, `r pre_article_code[1,"percent"]*100`% (*n* = `r pre_article_code[1,2]`) shared code at both stages. Relatively low code and data-sharing practices in our sample is consistent with analyses of sharing practices for published articles (e.g., [@o2021preferred]), even for journals with strict public data archiving policies [@roche2015public].

# Paving our future to open, transparent and community-driven science

Our analysis has allowed us to better understand preprinting/postprinting practices in *EcoEvoRxiv*. Overall, *EcoEvoRxiv* articles are diverse but with primary research articles on vertebrates comprising most of the articles posted. North America, Europe and Australia use *EcoEvoRxiv* the most with very few non-English language articles to date. Submitting authors who were earlier in their career and more often with ‘male-associated names’ tended to use  *EcoEvoRxiv* the most. Articles posted to *EcoEvoRxiv* tend to take up to 8 months to become published with many articles not being open access. Code and data sharing was also relatively uncommon at the preprint stage. At the same time, we attempted to collect data on community discussion around preprints no such data was found on preprint landing pages, likely reflecting inadequate functionality and cross-linking with sources where such discussion is occurring. Based on the insights from our analysis, we provide recommendations to authors and the scientific community on ways they can further promote open and transparent research through preprints:

- First, share your data and code at the preprint stage. Sharing data and code early can help improve the quality of research, establish precedence, and improve the transparency and computational reproducibility of scientific findings [@sarabipour2019value]. Reassuringly, sharing data and code is rarely associated with the 'scooping' of research findings [@soeharjono2021reported]. If authors are worried about data being used unintentionally, clear information surrounding its reuse can be included in a license (see https://choosealicense.com). Data can also be achived with an embargo on its reuse [@roche2014troubleshooting]. 
- Second, take advantage of peer-reviewing services such as Peer Community In (PCI). The time between posting a preprint and publication is still quite long (~8 months). One possible explanation is that preprints are not being sent to suitable journals or are struggling to get into review, slowing down constructive feedback that can improve the quality of a paper. Using PCI circumvents editorial decisions without review, yet only `r pci[2,3]*100`% (*n* = `r pci[2,2]`) used PCI. Using such services will ensure that authors receive faster feedback on a paper. Ninety-three journals currently accept PCI reviews and recommendations when considering a paper for publication (https://peercommunityin.org/pci-friendly-journals/).
- Third, seek out and contribute to constructive feedback on preprints [@bourne2017ten]. While it is clear that preprints help establish precedence and allow findings to be openly accessible, it still seems rare that constructive discussions form around preprints in an open forum (e.g. bioRxiv [@anderson2020biorxiv]). Unfortunately, the *EcoEvoRxiv* website does not provide opportunities for discussion given the limitations of the web server at this point in time. As such, we could not accurately assess how much discourse around a given preprint occurs. Clearly, as a community, we need to provide better platforms that document discussions around preprint findings. Such discussions help authors improve their work and communicate their findings more effectively (when done constructively, of course). One way to facilitate such discussions may be to use open preprint peer-review services such as *Peer Community In* (PCI) or *PubPeer* (see also [@avissar2024recommendations]) to provide feedback on preprints. 
- Finally, keep your preprints updated. While most preprints get seamlessly connected and merged with their published version, some remain 'disconnected' as separate articles. Incorrect cross-linking by indexing platforms (e.g., Google Scholar) can create confusion and lead to frustration among authors. Thankfully, the reasons for unmatched preprints and publications are well-understood and easily rectified. They often result from a mismatch between preprint and published metadata (e.g., titles and author details). For example, nearly one-third  of articles changed their title from preprint submission to publication [`r titles[2,3]*100`% (*n* = `r titles[2,2]`)]. We found that mismatched metadata almost always contributed to preprints and published articles not being matched automatically in Google Scholar. We recommend that authors update their preprints with the publication DOI when accepted to journals, especially if their title has changed. This is very easy for authors to do on *EcoEvoRxiv* and ensures that the preprint is correctly linked to the published article and citations are appropriately merged. Adding DOIs will also create a link between the final preprint and the published paper so that, no matter the open access status of the publication, the research findings will remain openly accessible, saving costs for authors (i.e., by not having to pay open access fees) and improving the visibility and use of research.

Despite the early successes of the new initiatives taken by *EcoEvoRxiv*, as described above, much work remains to be done to improve the understanding and use of pre- and postprints within our community. We view this perspective piece as a small step towards achieving that goal. We hope that readers are more familiar with the benefits of using community-driven preprint servers and the unique initiatives they can pursue. Community-driven preprint servers can set their own agenda and are driven by the needs and desires of the community. Supporting these initiatives should be a priority for all researchers. Volunteers at *EcoEvoRxiv* are encouraged to remain open to new and innovative ways to improve publication and open science practices. We believe that the future of preprints is bright, and community-driven initiatives, such as *EcoEvoRxiv* will play a crucial role in the future of scientific publishing.

# Acknowledgements
We would like to thank the California Digital Library (CDL) and the CDL team (particularly, Alainna Wrigley, Justin Gonder, Lisa Schiff, Catherine Mitchell, Hardy Pottinger and Amanda Karby) for their support in hosting and maintaining *EcoEvoRxiv* for the Society for Open, Reliable, and Transparent Ecology and Evolutionary Biology (SORTEE). We would like to thank Gabriela Hidalgo and Daisy Larios for helping connect us with the IUCN and facilitating discussions to make *EcoEvoRxiv* a place where IUCN reports can be posted. Finally, we would also like to thank the endless number of SORTEE volunteers, and those especially on the *EcoEvoRxiv* Committee, who have helped to make *EcoEvoRxiv* a success. This paper emerged from a hackathon at the 2023 SORTEE conference, and we thank delegates who attended the session but could not be part of this paper. DWAN would also like to thank the Australian Research Council for a Future Fellowship (FT220100276). SN and ML are supported by the Australian Research Council (ARC) Discovery Project Grants (DP210100812 and DP230101248).

# Data and Code Availability
All data and code can be found on GitHub at: https://github.com/daniel1noble/ecoevo_1000

# Conflict of Interest
The authors would like to acknowledge competing interests on the perspectives presented in this paper given that many (DWAN, SN, ML) are founding members of *EcoEvoRxiv* and/or are part of the *EcoEvoRxiv* committee.

# References