### 2024-09-26 proc_b/authors.qmd
Moving

---
title: ""
format:
  docx:
    reference-doc: ./bib/template.docx
---

Rafaela Almeida^1^, Kevin Bairos-Novak^2^, Heikel Balti^3^, Michael Bertram^4^, Louis Bliard^5^, Jack Brand^6^, Ilha Byrne^7^, Ying-Chi Chan^8^, Dena Clink^9^, Quentin Corbel^10^, Ricardo Correia^11^, Jordann Crawford-Ash^12^, Antica Culina^13^, Elvira D'Bastiani^14^, Gideon Gywa Deme^15^, Félicie Dhellemmes^16^, Shreya Dimri^17^, Szymek Drobniak^18^, Alexander Elsy^19^, Susan Everingham^20^, Samuel J. L. Gascoigne^21^, Matt Grainger^22^, Gavin Hossack^23^, Knut Anders Hovstad^24^, Ed Ivimey-Cook^25^, Matt Jones^26^, Ineta Kačergytė^27^, Georg Küstner^28^, Dalton Leibold^29^, Magdalena M. Mair^30^, Jake Martin^31^, April Robin Martinig^32^, Ayumi Mizuno^33^, Iain Moodie^34^, David Moreau^35^, Rose O'Dea^36^, James Orr^37^, Matthieu Paquet^38^, Rabindra Parajuli^39^, Joel Pick^40^, Patrice Pottier^41^, Marija Purgar^42^, Pablo Recio^43^, Dominique Roche^44^, Raphaël Royauté^45^, Saeed Shafiei Sabet^46^, Inês Silva^47^, Bruno E. Soares^48^, Birgit Szabo^49^, Alfredo Sánchez-Tójar^50^, Elina Takola^51^, Eli Thoré^52^, Bishnu Timilsina^53^, Wilco Verberk^54^, Stefan J.G. Vriend^55^, Kristoffer Wild^56^, Coralie Williams^57^, Nicholas Wu^58^, Zoe Xirocostas^59^, Yefeng Yang^60^, Melina de Souza Leite^61^, Natalie E. van Dis^62^, NA NA^63^

^1^KU Leuven, Belgium

^2^School of the Environment, University of Queensland, Australia

^3^UMR 6249 Chrono-Environnement, University of Bourgogne Franche-Comté, France

^4^Swedish University of Agricultural Sciences

^5^Department of Evolutionary Biology and Environmental Studies, University of Zurich

^6^Department of Wildlife, Fish, and Environmental Studies, Swedish University of Agricultural Sciences

^7^Scholf of the Environment, The University of Queensland

^8^Swiss Ornithological Institute, Switzerland

^9^K. Lisa Yang Center for Conservation Bioacoustics, Cornell University

^10^Station d'Ecologie Théorique et Experimentale, CNRS, France

^11^Biodiversity Unit, University of Turku, Finland

^12^Fenner School of Environment and Society, Australian National University, Canberra ACT

^13^Institute Ruder Boskovic, Zagreb. Croatia

^14^Department of Ecology and Evolutionary Biology, University of California, Los Angeles

^15^Case Western Reserve University

^16^Science of Intellingence, Berlin, Germany; Max Planck Institute for Human Development, Berlin, Germany

^17^Department of Evolutionary Biology, Bielefeld Univeristy, Germany

^18^UNSW/Jgaiellonian University

^19^ETH Zürich, Switzerland

^20^Institute of Plant Sciences and Oeschger Centre for Climate Change Research, University of Bern

^21^Department of Biology, University of Oxford, United Kingdom

^22^Norwegian Institute for Nature Research, Trondheim, Norway

^23^n/a

^24^Norwegian Biodiversity Information Centre, Trondheim, Norway

^25^University of Glasgow

^26^University of Exeter

^27^Ecology department, Swedish University of Agricultural sciences, Sweden

^28^Department of Animal Ecology and Tropical Biology, University of Würzburg, Germany

^29^Division of Ecology and Evolution, Research School of Biology, Australian National University

^30^Statistical Ecotoxicology, Bayreuth Center of Ecology and Environmental Research (BayCEER), University of Bayreuth, Germany

^31^Swedish university of agricultural sciences, Umeå, Sweden

^32^School of Biological, Earth and Environmental Sciences, UNSW, Sydney

^33^Department of Biology, Faculty of Science, Hokkaido University, Japan

^34^Department of Biology, Lund University, Sweden

^35^Centre for Brain Research, University of Auckland, New Zealand

^36^School of Agriculture, Food and Ecosystem Sciences, University of Melbourne

^37^Department of Biology, University of Oxford

^38^Station d'Ecologie Théorique et Experimentale, CNRS, France

^39^Department of Geosciences, Florida Atlantic University, Boca Raton, FL

^40^Institute of Ecology and Evolution, University of Edinburgh, UK

^41^School of Biological, Earth and Environmental Sciences, UNSW, Sydney

^42^Ruđer Bošković Institute, Zagreb, Croatia

^43^Department of Ecology and Evolution, RSB, Australian National University

^44^Department of Biology, Carleton University, Canada

^45^Université ParisSaclay, INRAE, AgroParisTech, UMR EcoSys, Palaiseau, France

^46^Fisheries Department, Faculty of Natural Sciences, University of Guilan

^47^Center for Advanced Systems Understanding, Görlitz, Germany

^48^University of Regina

^49^Division of Behavioural Ecology, University of Bern, Switzerland

^50^Department of Evolutionary Biology, Bielefeld University, Germany

^51^Helmholtz Center for Environmental Research (UFZ), Leipzig, Germany

^52^Department of Wildlife, Fish, and Environmental Studies, Swedish University of Agricultural Sciences

^53^Faculty of Biosciences, Fisheries and Economics, The Arctic University of Norway, Norway

^54^Radboud University Nijmegen

^55^Netherlands Institute of Ecology (NIOO-KNAW), Wageningen, Netherlands

^56^The University of Melbourne, School of BioSciences, Melbourne, VIC

^57^School of Biological, Earth and Environmental Sciences, UNSW, Sydney

^58^Hawkesbury Institute for the Environment, Western Sydney University

^59^School of Life Sciences, University of Technology Sydney

^60^School of Biological, Earth and Environmental Sciences, UNSW, Sydney

^61^Department of Ecology, University of São Paulo, Brazil

^62^Helsinki Institute of Life Sciences, Helsinki University, Finland

^63^NA